Introducing ICBe: An Event Extraction Dataset from Narratives about International Crises

### 2023-07-13 replication_paper/psrm_draft/ICBe_manuscript.tex
no keywords
doublespacing

\title{Introducing ICBe: Very High Recall and Precision Event Extraction
from Narratives about International Crises}

@@

\usepackage[utf8]{inputenc}
\usepackage{pifont}
\usepackage{newunicodechar}
\newunicodechar{✓}{\ding{51}}
\newunicodechar{✗}{\ding{55}}
\usepackage{array}
\usepackage{ctable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage[parfill]{parskip}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{cellspace}
\setlength\cellspacetoplimit{0.8ex}
\renewcommand{\arraystretch}{0.8}
\AtBeginEnvironment{lltable}{\singlespacing}
\usepackage{subcaption}
\usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
\usepackage{float}
\usepackage[algo2e]{algorithm2e}
\usepackage{changepage}
\renewcommand{\bibliography}[1]{}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\begin{document}
\maketitle

@@

\keywords{
    Diplomacy
   \and
    War
   \and
    Crises
   \and
    International Affairs
   \and
    Computational Social Science
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

If we could record every important interaction between countries in all
of diplomacy, military conflict, and international political economy,
how much unique information would this chronicle amount to, and how
surprised would we be to see something new? In other words, what is the
entropy of international relations? This record could in principle be
unbounded, but the central conceit of social science is that there are
structural regularities that limit what actors can do, their best
options, and even which actors are likely to survive (Brecher 1999;
Reiter 2015). If so, then these events can be systematically measured,
and accordingly, massive effort is expended in social science attempting
to record these regularities.\footnote{See work on crises (Brecher and
  Wilkenfeld 1982; Beardsley et al. 2020), militarized disputes (Palmer
  et al. 2021; Gibler 2018), wars (Sarkees and Wayman 2010; Reiter,
  Stam, and Horowitz 2016), organized violence (Ralph Sundberg and Mihai
  Croicu 2016; Davies, Pettersson, and Öberg 2022), political violence
  (Raleigh et al. 2010), sanctions (Felbermayr et al. 2020), and
  international agreements (Kinne 2020; Owsiak, Cuttner, and Buck 2018),
  dispute resolution (Frederick, Hensel, and Macaulay 2017), and
  diplomacy (Moyer, Turner, and Meisel 2020; Sechser 2011).} Thanks to
improvements in natural language processing, more open-ended efforts
have begun to capture entire unstructured streams of events from
international news reports.\footnote{See Beieler et al. (2016); Boschee
  et al. (2015); Brandt et al. (2018); Grant et al. (2017); Li et al.
  (2021). On event extraction from images and social media see Zhang and
  Pan (2019) and Steinert-Threlkeld (2019).} How close these efforts are
to accurately measuring all or even most of what is essential in
international relations is an open empirical question, one for which we
provide new evidence here.

Our contribution is a high coverage ontology and event dataset for key
historical episodes in 20th and 21st-century international relations. We
develop a large, flexible ontology of international events with the help
of both human coders and natural language processing. We apply it
sentence-by-sentence to an unusually high-quality corpus of historical
narratives of international crises (Brecher 1999; Wilkenfeld and Brecher
2000; Brecher et al. 2021). The result is a new lower bound estimate of
how much actually happens between states during pivotal historical
episodes. We then develop several methods for objectively gauging how
well these event codings reconstruct the information contained in the
original narrative. We conclude by benchmarking our event codings
against several current state of the art event data collection efforts.
We find that existing systems produce sequences of events that do not
contain enough information to reconstruct the underlying historical
episode. The underlying fine-grained variation in international affairs
is unrecognizable through the lens of current quantification efforts.

This is a measurement paper that makes the following argument --- there
is a real-world unobserved latent concept known as international
relations, we propose a method for systematically measuring it, we
successfully apply this method producing a new large-scale set of
measurements, and those measurements exhibit several desirable kinds of
internal and external validity, and those measurements out-perform other
existing approaches. The article organizes that argument into eight
sections: task definition; corpus; priors/existing state of the art;
ICBe coding process; internal consistency; case study selection; recall;
and precision. A final section concludes.

\hypertarget{task-definition}{%
\section{Task Definition}\label{task-definition}}

We consider the measurement task of abstracting discrete events about a
historical episode in international relations. The easiest way to convey
the task is with an example. Figure 1 shows a narrative account of the
Cuban Missile Crisis (1962) alongside a mapping from each natural
language sentence to discrete machine-readable abstractive events.
Formally, a historical episode, \(H\), is demarcated by a period of time
\([T_{start}, T_{end}] \in T\), a set of Players \(p \in P\), and a set
of behaviors they undertook during that time \(b \in B\). International
Relations, \(IR\), is the system of regularities that govern the
strategic interactions that world actors make during a historical
episode, given their available options, preferences, beliefs, and
expectations of choices made by others. We observe neither \(H\) nor
\(IR\) directly. Rather the Historical Record, \(HR\), produces
documents \(d \in D\) containing some relevant and true (as well as
irrelevant and untrue) information about behaviors that were undertaken
recorded in the form of unstructured natural language text. The task is
to combine informative priors about \(IR\) with an unstructured corpus
\(D\) to produce a series of structured discrete events, \(e \in E\),
that have high coverage, precision, and recall over what actually took
place in history, \(H\).

\begin{figure}[H]
\caption{Cuban Missile Crisis (1962) - ICB Narrative vs. ICBe Events \label{fig:case_study_cuban_precision}}
\centering{\includegraphics[height=19cm]{"./case_study_cuban_precision.png"}}
\end{figure}
\clearpage

\hypertarget{corpus}{%
\section{Corpus}\label{corpus}}

For our corpus, \(D\), we select a set of unusually high-quality
historical narratives from the International Crisis Behavior (ICB)
project (\(n=471\))(Supplementary Information (SI) Appendix 1.1)(Brecher
et al. 2021; Brecher and Wilkenfeld 1997). Their domain is 20th and
21st-century crises, defined as a change in the type, or an increase in
the intensity, of disruptive interaction with a heightened probability
of military hostilities that destabilizes states' relationships or
challenges the structure of the international system (Brecher and
Wilkenfeld 1982). Crises are a significant focus of detailed single case
studies and case comparisons because they provide an opportunity to
examine behaviors in IR short of, or at least prior to, full conflict
(Holsti 1965; Paige 1968; Allison and Zelikow 1971; Gavin 2014; Brecher
and Wilkenfeld 1982; Iakhnis and James 2019). Case selection was
exhaustive based on a survey of world news archives and region experts,
cross-checked against other databases of war and conflict, and
non-English sources (Kang and Lin 2019; Brecher et al. 2021, 59). Each
narrative was written by consensus by a small number of scholars, using
a uniform coding scheme, with similar specificity (Hewitt 2001). The
corpus is unique in IR because it is designed to be used in a downstream
quantitative coding project.

\hypertarget{prior-beliefs-about-ir-ontological-coverage-and-the-existing-state-of-the-art}{%
\section{Prior Beliefs about IR, Ontological Coverage, and the Existing
State of the
Art}\label{prior-beliefs-about-ir-ontological-coverage-and-the-existing-state-of-the-art}}

Next, we draw informative prior beliefs about the underlying process of
IR that we expect to govern behavior during historical episodes and
their conversion to the historical record. We organize our prior beliefs
along two overarching axes, summarized in detail by Table 1.

The first axis (rows) represents the types of information we expect to
find in IR and forms the basis for our proposed ontology. We employ a
metaphor of a chess game, with players (polities, rebel groups, IGOs,
etc.), pieces (military platforms, civilians, domains), and behaviors
(think, say, do). Precise sequencing is required to capture gambits
(sequences of moves) and outcomes (victory, defeat, peace, etc.), while
precise geo-coding is required to understand the chessboard (medium of
conflict). We find 472 actors and 117 different behaviors.\footnote{See
  the full codebook on Github Repository
  \href{https://urldefense.com/v3/__https://github.com/CenterForPeaceAndSecurityStudies/ICBEventData__;!!Mih3wA!WxDJtEczKfxGTh0S2Krunap8ReymFEL5iTWaSfOHeqlSdyfRx77zmjBSWO1OAm13$}{ICBEventData}.}

We base our informed priors primarily on two sources of information. The
first is the extensive existing measurement efforts of IR which we
provide citations alongside each concept. Second, we performed
preliminary natural language processing of the corpus and identified
named entities and behaviors mentioned in the text. Verbs were matched
to the most likely definition found in Wordnet (Miller 1995), tallied,
and then aggregated into a smaller number of hypernyms balancing
conceptual detail and manageable sparsity for human coding (SI Appendix
1.2).

The second axis (columns) compares the very high ontological coverage of
ICBe to existing state of the art systems in production and with global
coverage. They begin with our contribution ICBe, alongside other
event-level datasets including CAMEO dictionary lookup-based systems
(Historical Phoenix (Althaus et al. 2019); ICEWS (Boschee et al. 2015);
Terrier (Grant et al. 2017)), the Militarized Interstate Disputes
Incidents dataset, and the UCDP-GED dataset (Davies, Pettersson, and
Öberg 2022; Sundberg and Melander 2013).\footnote{Other related datasets
  that insufficiently overlap ICBe's domain for comparison include BCOW
  (Leng and Singer 1988), WEIS (McClelland 1978), CREON (Hermann 1984),
  CASCON (Bloomfield and Moulton 1989), SHERFACS (Sherman 2000),
  Real-Time Phoenix (Brandt et al. 2018), and COfEE (Balali, Asadpour,
  and Jafari 2021) (see histories in Merritt (1994) and Schrodt and Hall
  (2006)).} The final set of columns compares episode-level datasets
beginning with the original ICB project (Brecher et al. 2021; Brecher
and Wilkenfeld 1982; Beardsley et al. 2020); the Militarized Interstate
Disputes dataset (Palmer et al. 2021; Gibler 2018), and the Correlates
of War (Sarkees and Wayman 2010). With the exception of large-scale
CAMEO dictionary-based systems, the existing state of the art
quantitative datasets ignore the vast majority of the information
content found in international relations.\footnote{See Balali, Asadpour,
  and Jafari (2021) for a recent review of ontological depth and
  availability of Gold Standard example text.}

\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}

\setlength{\tabcolsep}{2pt}

@@

@@

\caption{Ontological coverage of ICBe versus existing State of the Art
}\\

@@

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}p{\dimexpr 0.1in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{Concept}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.6in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{Literature}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}p{\dimexpr 0.22in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{ICBe}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.22in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{Phoenix}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.22in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{Terrier}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.22in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{ICEWs}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.22in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{MID\ Incidents}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}p{\dimexpr 0.22in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{UCDP-GED}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.22in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{ICB}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.22in+0\tabcolsep+0\arrayrulewidth}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{MIDs}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.22in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 2pt}}{\rotatebox[origin=c]{270}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{COW}}}} \\

@@

\endfirsthead

>>>>

\title{Introducing ICBe: An Event Extraction Dataset from Narratives
about International Crises}

@@

\usepackage[utf8]{inputenc}
\usepackage{pifont}
\usepackage{newunicodechar}
\newunicodechar{✓}{\ding{51}}
\newunicodechar{✗}{\ding{55}}
\usepackage{array}
\usepackage{ctable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage[parfill]{parskip}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{cellspace}
\setlength\cellspacetoplimit{0.8ex}
\renewcommand{\arraystretch}{0.8}
\AtBeginEnvironment{lltable}{\singlespacing}
\usepackage{subcaption}
\usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
\usepackage{float}
\usepackage[algo2e]{algorithm2e}
\usepackage{changepage}
\renewcommand{\bibliography}[1]{}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\begin{document}
\maketitle

@@

\doublespacing

If we could record every international interaction in the realms of
diplomacy, conflict, economics, and beyond, how much unique information
would this chronicle amount to, and how surprised would we be to see
something new? In other words, what is the entropy of international
relations? While this record could in principle be unbounded, the
central conceit of social science is that there are structural
regularities that limit what actors can do, their best options, and even
which actors are likely to survive
(\textbf{brecherInternationalStudiesTwentieth1999?};
\textbf{reiterShouldWeLeave2015?}). If so, then these events can be
recorded and systematically measured by social scientists interested in
these regularities.\footnote{See work on crises
  (\textbf{brecherCrisesWorldPolitics1982?};
  \textbf{beardsleyInternationalCrisisBehavior2020?}), militarized
  disputes (\textbf{palmerMID5Dataset20112021?};
  \textbf{giblerInternationalConflicts181620102018?}), wars
  (\textbf{sarkeesResortWar181620072010?};
  \textbf{reiterRevisedLookInterstate2016?}), organized violence
  (\textbf{ralphsundbergUCDPGEDCodebook2016?};
  \textbf{daviesOrganizedViolence19892022?}), political violence
  (\textbf{raleighIntroducingACLEDArmed2010?}), sanctions
  (\textbf{felbermayrGlobalSanctionsData2020?}), and international
  agreements (\textbf{kinneDefenseCooperationAgreement2020?};
  \textbf{owsiakInternationalBorderAgreements2018?}), dispute resolution
  (\textbf{frederickIssueCorrelatesWar2017?}), and diplomacy
  (\textbf{moyerWhatAreDrivers2020?};
  \textbf{sechserMilitarizedCompellentThreats2011?}).} Thanks to
improvements in natural language processing, more open-ended efforts
have begun to capture entire unstructured streams of events from
international news reports.\footnote{See
  (\textbf{beielerGeneratingPoliticalEvent2016?});
  (\textbf{boscheeICEWSCodedEvent2015?});
  (\textbf{brandtPhoenixRealTimeEvent2018?});
  (\textbf{grantOUEventData2017?});
  (\textbf{liComprehensiveSurveySchemabased2021?}). On event extraction
  from images and social media see
  (\textbf{zhangCASMDeepLearningApproach2019?}) and
  (\textbf{steinert-threlkeldFutureEventData2019?}).} This has invited
fruitful efforts to evaluate the coverage, quality, and accuracy of
attempts to measure international affairs.

We advance existing efforts to identify and structure regularized events
and actor in international politics by combining human coding with
natural language processing to develop (1) a large, flexible ontology of
international affairs and (2) a fine-grained and structured event
dataset of international crises from 1918-2017 developed by applying our
ontology to an unusually high-quality corpus of historical narratives of
international crises
(\textbf{brecherInternationalStudiesTwentieth1999?};
\textbf{wilkenfeldInterstateCrisesViolence2000?};
\textbf{brecherInternationalCrisisBehavior2021?}). We then develop
several methods for objectively gauging how well these event codings
reconstruct the information contained in the original crisis narrative.
We conclude by benchmarking our event codings against several current
state of the art event data collection efforts. The underlying
fine-grained variation in international affairs is unrecognizable
through the lens of current quantification efforts. We find that
existing models produce data on historical episodes that do not contain
enough information to reconstruct the underlying event. In focusing this
initial effort on international crises as a proof of concept sample, we
demonstrate our ontology and method's potential to improve upon existing
empirical identifications of patterns of international interactions.

In the proceeding four sections, this measurement paper makes the
following arguments. First, there is a real-world unobserved latent
concept known as international relations that can and should be
systematically measured. Second, we propose a method for systematic
large-scale measurement of the actors and behaviors in international
affairs and as a proof of concept apply that method to a well-regarded
and salient sample of events known as international crises. Third, in
doing so we confirm that those measurements exhibit several desirable
kinds of internal and external validity and out-perform existing
approaches. Fourth, this validation can be evaluated in detail via new
event visualizations, with examples provided for case studies of the
1962 Cuban Missile Crisis and 2014 Crimea-Donbas crisis. A final section
concludes.

\hypertarget{identifying-and-measuring-international-relations}{%
\section{Identifying and measuring international
relations}\label{identifying-and-measuring-international-relations}}

\hypertarget{motivation}{%
\subsection{Motivation}\label{motivation}}

How can scholars abstract and measure discrete events about a historical
episode in international relations? We employ a metaphor of chess, a
game that despite its complexity can nonetheless be recorded in a
standardized and structured manner. Like chess, international
interactions involve a finite set of players (e.g.~polities, rebel
groups, IGOs) expressing their behaviors (e.g.~thinking, saying, and
doing) by moving pieces (e.g.~military platforms, civilian personnel,
diplomats) from and onto identifiable locations (geo-coded coordinates).
These moves occur over a marked time period (start and end of a
historical episode) and gambits occur in recordable sequences (actions
and reactions) to produce observable, if still disputable, outcomes
(e.g.~victory, defeat, stalemate, peace). Our knowledge of this
historical episode, including the actors involved as well as their
preferences, behaviors, and beliefs are only indirectly observed from
historical records that most often take the form of unstructured natural
language text.

Much like the recording of chess evolved from natural language
descriptive text notation to the modern figurine algebra notion,
international relations scholars have recently sought to produce a
structured account of historical events by combining the unstructured
corpus of historical records with informative priors about international
relations. The resulting structured account of information about a
historical episode can be combined with that of other events to produce
a systematic account of political events from which we can garner novel
insight, assuming the underlying data have high coverage, precision, and
recall. The easiest way to convey the desired produced of this task is
with an example. (\textbf{fig-cuba-narrative?}) shows a narrative
account of the Cuban Missile Crisis (1962) in natural language sentences
alongside a mapping to discrete machine-readable abstractive events.
From the structured data, scholars can identify similarities and
differences across events concerning important concepts like when
particular foreign policy actions deter versus inflame
(\textbf{jervisCooperationSecurityDilemma1978?};
\textbf{glaserCausesConsequencesArms2000?}), when third parties mediate
in interstate disputes
(\textbf{haffarEmergentPeacemakersCataloguing2002?};
\textbf{quinnPowerPlayMediation2006?}), and how actors try to
communicate resolve (\textbf{tragerDiplomacyWarPeace2016?};
\textbf{luptonReexaminingReputationResolve2018?}). Identifying patterns
of international interactions is not just an inherently interesting
enterprise; it is a necessary precondition to important efforts to
predict where policymakers should turn their attention to improve global
welfare (\textbf{wardLearningSteppingFuture2013?};
\textbf{begerReassessingRoleTheory2021?}).

\begin{figure}
\hypertarget{fig-cuba-narrative}{%
\centering
\includegraphics{case_study_cuban_precision.png}
\caption{Cuban Missile Crisis (1962) - ICB Narrative vs.~ICBe
Events}\label{fig-cuba-narrative}
}
\end{figure}

\hypertarget{existing-state-of-the-art-measurements}{%
\subsection{Existing state of the art
measurements}\label{existing-state-of-the-art-measurements}}

We draw informative prior beliefs about the underlying process of
international relations that we expect to govern behavior during
historical episodes and their conversion to the historical record. We
organize our prior beliefs along two overarching axes: existing efforts
to identify the actors/actions of international relations and
identifying a corpus that can be used to produce an ontology of the
information we hope to recover. Table 1 describes these two axes as
columns and rows, respectively.

As we are not the first to attempt to measure international relations in
a structured manner, the columns of Table 1 compare the ontological
coverage of ICBe to existing state of the art systems in production and
with global coverage. We choose these datasets and models as they
represent frequently used and reputable efforts to structure and
describe historical events of interest to scholars of international
politics. The first column starts with our contribution, ICBe, alongside
other event-level datasets including CAMEO dictionary lookup-based
systems (Historical Phoenix
(\textbf{althausClineCenterHistorical2019?}); ICEWS
(\textbf{boscheeICEWSCodedEvent2015?}); Terrier
(\textbf{grantOUEventData2017?})), the Militarized Interstate Disputes
Incidents dataset, and the UCDP-GED dataset
(\textbf{daviesOrganizedViolence19892022?};
\textbf{sundbergIntroducingUCDPGeoreferenced2013?}).\footnote{Other
  related datasets that insufficiently overlap ICBe's domain for
  comparison include BCOW
  (\textbf{lengMilitarizedInterstateCrises1988?}), WEIS
  (\textbf{mcclellandWorldEventInteraction1978?}), CREON
  (\textbf{hermannComparativeResearchEvents1984?}), CASCON
  (\textbf{bloomfieldCASCONIIIComputeraided1989?}), SHERFACS
  (\textbf{shermanSHERFACSCrossParadigmHierarchical2000?}), Real-Time
  Phoenix (\textbf{brandtPhoenixRealTimeEvent2018?}), and COfEE
  (\textbf{balaliCOfEEComprehensiveOntology2021?}) (see histories in
  (\textbf{merrittMeasuringEventsInternational1994?}) and
  (\textbf{schrodtTwentyYearsKansas2006?})).} The final set of columns
compares episode-level datasets beginning with the original ICB project
(\textbf{brecherInternationalCrisisBehavior2021?};
\textbf{brecherCrisesWorldPolitics1982?};
\textbf{beardsleyInternationalCrisisBehavior2020?}); the Militarized
Interstate Disputes dataset (\textbf{palmerMID5Dataset20112021?};
\textbf{giblerInternationalConflicts181620102018?}), and the Correlates
of War (\textbf{sarkeesResortWar181620072010?}). There is imperfect
overlap concerning their intended depth and scope of coverage;
`international crises' are similar, but not identical to, `interstate
wars' and `militarized interstate disputes' which differ yet again from
`individual events of organized violence' and `non-violent action'. Even
like-concepts require care in comparison, as an `aim' in ICBe is the
same as in MIPS, but an `alert' in ICBe is not the same as an `alert' in
MID. Definitions for each ICBe variable are provided in the codebook.

\clearpage
\singlespacing

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{0pt}

@@

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

@@

\caption{Ontological\ coverage\ of\ ICBe\ versus\ existing\ State\ of\ the\ Art}\\

@@

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}m{\dimexpr 0.1in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 2in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Concept}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.6in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Literature}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{ICBe}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Phoenix}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Terrier}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{ICEWs}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{MID\ Incidents}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{UCDP-GED}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{ICB}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{MIDs}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}!{\color[HTML]{000000}\vrule width 2pt}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{COW}}}} \\

\noalign{\global\arrayrulewidth 2pt}\arrayrulecolor[HTML]{000000}

\hhline{|>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}-|>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}-|>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}-}\endfirsthead \caption[]{Ontological\ coverage\ of\ ICBe\ versus\ existing\ State\ of\ the\ Art}\\

@@

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}m{\dimexpr 0.1in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 2in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Concept}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.6in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Literature}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{ICBe}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Phoenix}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Terrier}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{ICEWs}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{MID\ Incidents}}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{UCDP-GED}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{ICB}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{MIDs}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}!{\color[HTML]{000000}\vrule width 2pt}}{\rotatebox[origin=c]{270}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{COW}}}} \\

\noalign{\global\arrayrulewidth 2pt}\arrayrulecolor[HTML]{000000}

\hhline{|>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}-|>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}-|>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=2pt}-}\endhead
